---
title: "Similation od Distributed Averages of 40 iid exponentials"
author: "Nandini Guntur"
date: "5/11/2020"
output: pdf_document
---

# Statistical Inference Course Project
Part I - A Simulation Exercise
## Overview :
In this project we will be investigating the exponential distribution in R and compare it with the Central Limit Theoram. The exponential distribution can be simulated in R with rexp(n,lambda) where lambda is the rate parameter. The mean and standard deviation of the exponential distribution is 1/lambda. We will set lambda = 0.2 for all the simulations. We will investigate the distribution of 40 iid exponentials and do a thousand simulations.
 
Setting a seed for reproducibility :

```{r, echo=TRUE}
set.seed(111)
```
Setting a rate parameter lambda :

```{r, echo=TRUE}
lambda = 0.2
```
Setting a size of a sample :

```{r, echo=TRUE}
n = 40
```
Setting the number of simulations :

```{r, echo=TRUE}
nosim = 1000
```
Generating 10000 samples of 40 exponentials and calculating their mean values "

```{r, echo=TRUE}
r <- replicate(nosim, rexp(n,lambda))
dim(r)
```
We can see that r is a matrix of 40 rows and 1000 columns. Since each column contains a sample of 40 exponentials, we will apply mean() to columns to obtain 1000 sample means.

```{r, echo=TRUE}
exp_means <- apply(r, 2, mean)
```
### 1. Show the sample mean and compare it to the theoritrical mean of the distribution
Sample(empirical) Mean :

```{r, echo=TRUE}
e_mean <- mean(exp_means)
e_mean
```
Theorical Mean :

```{r, echo=TRUE}
t_mean <- 1/lambda
t_mean
```
We can see that the sample mean 5.02562 is a good approximation of the theorical mean od 5.

```{r, echo=TRUE}
means <- cumsum(exp_means)/(1:nosim)
library(ggplot2)
g <- ggplot(data.frame(x = 1: nosim , y=means), aes(x = x, y = y))
g <- g+geom_hline(yintercept <- t_mean, color = "red")+geom_line(size=1)
g <- g+labs(x = "Number of Simulations", y = "Cummulative Mean")
g <- g+ggtitle("Sample Mean of 1000 Samples of 40 Random Exponentials")
g
```
As we can see from the graph, empirical mean is the consistent estimator of the theotrical mean, because it converges to the value of theotrical mean.

### 2. Show how variable the sample is (via variance) and compare it to the theotrical variance of the distribution.
Sample Variance :

```{r, echo=TRUE}
e_var <- var(exp_means)
e_var
```
Theotrical Variance of the distribution of sample means 

```{r, echo=TRUE}
t_var <- (1/(lambda*sqrt(n)))^2
t_var
```

As we can see from the graph, both empirical and theotrical variance of the distribution of the sample means have value close to 0.6

```{r, echo=TRUE}
cumvar <- cumsum(exp_means - e_means)^2/seq_along(exp_means - 1))
g <- g+geom_hline(yintercept <- t_var, color = "red")+geom_line(size=1)
g <- g+labs(x = "Number of Simulations", y = "Cummulative Mean")
g <- g+ggtitle("Sample Mean of 1000 Samples of 40 Random Exponentials")
g
```
As we can see from the graph, sample variance is a consistant estimator of the theotrical variance, because it converges to the value of the theotrical variance.

### 3. Show the distribution is approximately normal.

```{r, echo=TRUE}
g <- ggplot(data.frame(x = exp_means), aes(x = x))
g <- g + geom_histogram(aes(y=...density...), color = "white", fill= "salmon")
g <- g + stat_function(fun = "dnorm", color = "black" , args = list(mean = t_mean, sd = se))
g <- g+geom_vline(xintercept <- t_mean, color = "black" , size =1)
g <- g+ggtitle("Distribution of Average of 1000 Samples of 40 Random Exponential Variables")
g <- g+xlabs("Means")
g <- g+ylabs("Density")
g
```
As we can see from the graph, distribution of average of 1000 samples of 40 random exponential variable is normal.
